---
title: "EDA_FINAL_CODE"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Dataset
```{r, echo=TRUE}
data <- read.csv('HR_Employee_Data.csv')
```

# Import Libraries
```{r, echo=TRUE}
library(tidyverse)
library(plotly)
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)
library(highcharter)
library(randomForest)
library(e1071)
library(caret)
library(kernlab)
library(caTools)

library(plotly)
library(corrly)
library(ecodist)
library(naivebayes)
library(psych)
library(rpart.plot)
library(ggcorrplot)

library(ROCR) 
library(class)
#install.packages("ggcorrplot")
#install.packages("ROCR")
#install.packages("rpart.plot")
#install.packages('ecodist')
#install.packages("remotes")
#remotes::install_github("kmaheshkulkarni/corrly")
#install.packages("hrbrthemes")
#install.packages("highcharter")
#install.packages("kernlab")
#install.packages("caTools")

```

# Summary of the Data
```{r, echo=TRUE}
head(data)
str(data)
#glimpse(data)
summary(data)
```


# Check for Null Values
```{r, echo=TRUE}
cbind(lapply(lapply(data, is.na), sum))
sum(is.na(data))
```

# Data Cleaning
```{r}
data$satisfaction_level<-gsub("%","",as.character(data$satisfaction_level))
data$satisfaction_level=as.integer(data$satisfaction_level)
head(data)
```


```{r, echo=TRUE}
data$last_evaluation<-gsub("%","",as.character(data$last_evaluation))
data$last_evaluation=as.integer(data$last_evaluation)
head(data)
```

#Renaming the Column name in Dataframe
```{r,echo=TRUE}
data <- data %>%
  rename(Emp_Id=ï..Emp_Id )
head(data)

```

# 1)Correlation plot
```{r, echo=TRUE}

corr <- round(cor(data[2:9]), 1)
ggcorrplot(corr, lab = TRUE)

```

# 2) People who have left in each department
```{r, echo=TRUE}

ans=crosstab(data$Department,data$left)
Department=rownames(ans)
fig <- plot_ly(ans,x = ~Department, y = ~X0, type = 'bar', name = 'Working for Company')
fig<- fig %>% add_trace(y =~X1, name = 'Left the Company')
fig <- fig %>% layout(title="Employees who have left based on department", yaxis = list(title = 'Count'), barmode = 'group')
fig
```

# 3) People who have left based on salary
```{r, echo=TRUE}

ans=crosstab(data$salary,data$left)
Salary=rownames(ans)
fig <- plot_ly(ans,x = ~Salary, y = ~X0, type = 'bar', name = 'Working for Company')
fig<- fig %>% add_trace(y =~X1, name = 'Left the Company')
fig <- fig %>% layout(title="Employees who have left based on salary", yaxis = list(title = 'Count'), barmode = 'group')
fig
```

# 4) Area plot of time spent in company compared with those who have nand have not left 
```{r, echo=TRUE}
ans=crosstab(data$time_spend_company,data$left)
ans
Time_Spent=rownames(ans)
Time_Spent
fig <- plot_ly(ans,x = ~Time_Spent, y = ~X0, type = 'scatter', mode = 'lines', name = 'Working for Company', fill = 'tozeroy')
fig <- fig %>% add_trace(y = ~X1, name = 'Left the Company', fill = 'tozeroy')
fig <- fig %>% layout(xaxis = list(title = 'Time Worked'),
         yaxis = list(title = 'Count'))
fig
```

# 5) Average monthly working hours according to department
```{r, error=TRUE}
df <- data
head(df)
str(df)
data1 <- df%>%
  group_by(Department)%>%
  summarize(Avg_hrs = mean(average_montly_hours))

fig <- plot_ly(data1, x = ~Department, y = ~Avg_hrs, type = 'bar', color = I("dark blue"))
fig <- fig %>% layout(title = "Average monthly working hours according to department",
         xaxis = list(title = "Department"),
         yaxis = list(title = "Average monthly working hours"))
fig
#ggplot(df,aes(x=number_project,y=average_montly_hours))+geom_jitter(aes(color=Department))
```

## 6) Number of work accidents for each department
```{r, error=TRUE}
data2 <- df%>%
  filter(Work_accident==1)%>%
  group_by(Department)%>%
  summarize(No_of_wa = n())%>%
  arrange(No_of_wa)
head(data2)
hc <- data2 %>% 
  hchart('line', hcaes(x = Department, y = No_of_wa))%>%
  hc_title(text = "Number of work accidents for each department")%>%
  hc_yAxis(title = "Number of work accidents")
hc
```

## 7) Density plot of satisfaction level according to salary
```{r, error=TRUE}

l <- df %>% filter(salary == "low")
m <- df %>% filter(salary == "medium")
h <- df %>% filter(salary == "high")
hc2 <- hchart(
  density(l$satisfaction_level), type = "area", 
  color = "steelblue", name = "Low Salary"
  ) %>%
  hc_add_series(
    density(m$satisfaction_level), type = "area",
    color = "#B71C1C", 
    name = "Medium Salary"
    )%>%
  hc_add_series(
    density(h$satisfaction_level), type = "area",
    color = "yellow", 
    name = "High Salary"
    )%>%
  hc_title(text = "Density plot of satisfaction level according to salary")%>%
  hc_xAxis(title = "Satisfaction Level (0-100)")
hc2
```

## 8) Time spent per Department
```{r, error=TRUE}

fig <- plot_ly(df, labels = ~Department, values = ~time_spend_company, type = 'pie')
fig <- fig %>% layout(title = 'Time spent per Department',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig
```

## 9) Distribution of Satisfaction level

```{r, echo=TRUE}
  p <- ggplot(data, aes(x =satisfaction_level))+
      geom_bar(color="darkblue", fill="lightblue")+
      ggtitle("Distribution of Satisfaction level") +
      theme(axis.text.x = element_text(angle = 90, hjust = 1))
    p
```

## 10) Number of projects for each department
```{r, echo=TRUE}
fig <- plot_ly(data,x=~Department, y=~number_project,color = ~salary,type="bar")
fig
```

## 11) Box plot between satisfaction_level and average_montly_hours.
```{r, echo=TRUE}
data %>%
  ggplot(aes(x=satisfaction_level,y=average_montly_hours))+
  geom_boxplot(fill="lightblue")+
  xlab("satisfaction_level")+
  ylab("average_montly_hours")+
  facet_grid(~salary)
```

## 12) Salary distribution
```{r, echo=TRUE}
fig <- plot_ly(data,labels = ~salary, values = ~time_spend_company,type="pie", textinfo='label+percent')
fig
```

## 13) Promotion in last 5 years vs number of projects
```{r, echo=TRUE}
fig <- plot_ly(data, x = ~promotion_last_5years, y = ~number_project ,type = 'bar', color=~Department)
fig
```

# ML MODELS:

## 1)DT MODEL
```{r}

set.seed(234)
dataDT=data[2:11]
smpl<-sample(2,nrow(dataDT),replace=T,prob=c(0.8,0.2))
train<-dataDT[smpl==1,]
test<-dataDT[smpl==2, ]
fit <- rpart(left~., data = train, method = 'class')
rpart.plot(fit, extra = 106)
```

## Prediction
```{r}
predict_unseen <-predict(fit, test, type = 'class')
table_mat <- table(test$left, predict_unseen)
table_mat
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test
```

### Accuracy: [1] 0.9720093

## 2)Naive Bayes
```{r, echo=TRUE}

data$left<-as.factor(data$left)
#data$rank<-as.factor(data$rank)
str(data)
dataNB<-data[2:11]
#pairs.panels(data) 
#cor(data$gre,data$gpa)
set.seed(234)
smpl<-sample(2,nrow(dataNB),replace=T,prob=c(0.8,0.2))
train<-dataNB[smpl==1,]
test<-dataNB[smpl==2, ]
mdl<-naive_bayes(left~ .,data=train)
#mdl
```

## NB Prediction
```{r, echo=TRUE}
#plot(mdl)
p<-predict(mdl,train,type='prob')
head(cbind(p,train))
p1<-predict(mdl,train)
(tab1<-table(p1,train$left))
accuracy=sum(diag(tab1))/sum(tab1)
accuracy
```

### Accuracy : [1] 0.7914652


## 3)SVM Model
```{r, error=TRUE}
df2 <- df
df2$left <- as.factor(df2$left)
str(df2)
df2$Emp_Id <- NULL
set.seed(234)
split <- sample.split(df2, SplitRatio = 0.7)
split
train <- subset(df2, split == "TRUE")
test <- subset(df2, split == "FALSE")
classifier = svm(formula = left ~ .,
                 data = train,
                 type = 'C-classification',
                 kernel = 'linear')
y_pred = predict(classifier, newdata = test[-7])
y_train_pred = predict(classifier, newdata = train[-7])
cm = table(test[, 7], y_pred)
cm
cm2 = table(train[, 7], y_train_pred )
cm2
sum(diag(cm))/sum(cm)
```

## Accuracy: [1] 0.7751111

## 4)Random Forest Model
```{r, error=TRUE}
# Splitting data in train and test data
# Fitting Random Forest to the train dataset
set.seed(120)  # Setting seed
classifier_RF = randomForest(x = train[-7],
                             y = train$left,
                             ntree = 50)
classifier_RF
# Predicting the Test set results
y_pred = predict(classifier_RF, newdata = test[-7])
# Confusion Matrix
confusion_mtx = table(test[, 7], y_pred)
confusion_mtx
# Plotting model
plot(classifier_RF)
# Importance plot
importance(classifier_RF)
# Variable importance plot
varImpPlot(classifier_RF)

sum(diag(confusion_mtx))/sum(confusion_mtx)
```

### Accuracy: [1] 0.9888889

## 5)Logistic Regression
```{r, echo=TRUE}
# Loading package


set.seed(234)   
# Splitting dataset
split <- sample.split(data, SplitRatio = 0.8)
split
   
train<- subset(data[2:11], split == "TRUE")
test<- subset(data[2:11], split == "FALSE")
   
# Training model
logistic_model <- glm(left ~ satisfaction_level + last_evaluation, data = train, family = "binomial")
logistic_model
   
# Summary
summary(logistic_model)
   
# Predict test data based on model
predict<- predict(logistic_model,test, type = "response")
#predict  

# Changing probabilities
predict<- ifelse(predict>0.5, 1, 0)
   
# Evaluating model accuracy
# using confusion matrix
table(test$left, predict)
   
missing_classerr <- mean(predict != test$left)
print(paste('Accuracy =', 1 - missing_classerr))
   
# ROC-AUC Curve
ROCPred <- prediction(predict, test$left) 
ROCPer <- performance(ROCPred, measure = "tpr", x.measure = "fpr")
   
auc <- performance(ROCPred, measure = "auc")
auc <- auc@y.values[[1]]
auc
   
# Plotting curve
plot(ROCPer)
plot(ROCPer, colorize = TRUE,print.cutoffs.at = seq(0.1, by = 0.1),main = "ROC CURVE")
abline(a = 0, b = 1)
   
auc <- round(auc, 4)
legend(.6, .4, auc, title = "AUC", cex = 1)

```

### Accuracy :  0.771149144254279

## 6)KNN
```{r, echo=TRUE}
f=data$left
data = subset(data, select = -c(left) )
data$left <- f
data_df<-data
data$Department= as.numeric(as.factor(data$Department))
data$salary= as.numeric(as.factor(data$salary))
head(data_df)
set.seed(123)
split <- sample.split(data, SplitRatio = 0.7)
train_cl <- subset(data, split == "TRUE")
test_cl <- subset(data, split == "FALSE")
  
# Feature Scaling
train<- scale(train_cl[, 2:10])
test<- scale(test_cl[, 2:10])


# Fitting KNN Model 
# to training dataset
# K = 3
classifier_knn <- knn(train = train,
                      test = test,
                      cl = train_cl$left,
                      k = 3)

# Confusiin Matrix
cm <- table(test_cl$left, classifier_knn)
cm

# Model Evaluation - Choosing K
# Calculate out of Sample error
misClassError <- mean(classifier_knn != test_cl$left)
print(paste('Accuracy =', 1-misClassError))
 
```

### Accuracy: 0.949761


### KNN prediction
```{r, echo=TRUE}
#head(data)
satisfaction_level<-57
last_evaluation<-34
number_project<-3
average_montly_hours<-130
time_spend_company<-3
Work_accident<-0
promotion_last_5years<-1

s1<-levels(factor(data_df$Department))
s1
s2<-as.numeric(levels(factor(data$Department)))
s2
Department<-"IT"
p<-as.numeric(match(Department,s1))
Department=s2[p]
Department


a1<-levels(factor(data_df$salary))
a1
a2<-as.numeric(levels(factor(data$salary)))
a2
salary<-"medium"
p1<-as.numeric(match(salary,a1))
salary=a2[p1]
salary

l1<-levels(factor(data_df$left))
l1
l2<-as.numeric(levels(factor(data$left)))
l2

x=c(satisfaction_level,last_evaluation,number_project,average_montly_hours,time_spend_company,Work_accident,promotion_last_5years,Department,salary)

classifier_knn <- knn(train = train,
                      test = test,
                      cl = train_cl$left,
                      k = 3)


z<-knn(train=train_cl[,2:10],test = x,cl = train_cl$left, k = 3)
z1<-as.numeric(match(z,l2))
cat("Left :",l2[z1])
```


